---
title: "Homework-2"
author: "Shiying Wang"
date: "10/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Effect on the out-of-sample mean square error as λ varies in ridge regression

Firstly, we create a predict.ridge function to get predicted values of test_ridge dataset based on the parameters obtained from the ridge regression on train_ridge dataset.
```{r }
library(bis557)
data(ridge_train)
data(ridge_test)
predict.ridge<-function(object,...)
{
  data<-list(...)[[1]]
  mm<-model.matrix(object$form,data)
  mm %*% object$coefficients
}
```

Then,we use a series of lambda values to fit ridge regression model on ridge_train dataset, and calculate out-of-sample mean square error.
```{r}
MSE<-rep(NA,50)
for(i in 1:50)
{
train_fit<-ridge_reg(y~.,i,ridge_train)
error<-ridge_test$y-predict.ridge(train_fit,ridge_test)
MSE[i]<-mean(error^2)
}
```
Finally, we draw the plot of MSE as lambda varies.
```{r}
plot(MSE)
```
  
When $0<$lambda$<30$, MSE decreases as lambda increases. When lambda$>30$, MSE increases as lambda increases. MSE reaches its minimum value when lambda is around 30.  


In conclusion, there is a certain value of lambda, so that out-of-sample mean square error reaches its minimum.


